---
title: "QMB Exercise 2 - Estimation and Testing"
author: "Karin Gryzlak & Dino Nienhold"
date: "10 May, 2015"
output:
  pdf_document:
    fig_caption: yes
    
---

#Introduction
The following report is based on the QMB Exercise 2 - Estimation and Testing. The task description pdf file is bis_ex2-EstimationTesting-20150429.pdf

#Requirements
Please make sure that you the following packages loaded in your workspace.
```{r, message=F, warning=F}
library("dplyr")
library("ggplot2")
library("ggExtra")
library("gridExtra")
```

#Data Set
Please make sure you have the file housingrents.csv in the subdirectoy Data in your workspace.
```{r}
housingrents <- read.csv("./Data/housingrents.csv",sep=";")
```
\pagebreak

##Data Processing
For analysis purposes it is necessary to convert the rooms and NRE variable to a factor. Furthermore a new variable rps (rent per square meter) is created
```{r}
housingrents <- mutate(housingrents, rooms = factor(rooms), 
  nre = factor(nre,levels=c(0,1),labels=c("no","yes")))
housingrents <- mutate(housingrents,rps = rent/area)
```

##Task 1
...
```{r}

```

##Conclusion
...


\pagebreak

##Task 2
Task 2 checks normality of variable rent per square (rps). Additionally t-tests are conducted

### a)
In this sub task the normal distribution of the variable rpf the NRE respectively non-NRE appartments are checked.
Create two datasets as a first step:
```{r}
nrehousing <- filter(housingrents,nre=="yes")
nonnrehousing <- filter(housingrents,nre=="no")
```
Check normal distribution of rps variable for nre houses:

p1 <- qplot(x = 1, y = rps, data = nrehousing, xlab = "", geom = 'boxplot') +
        coord_flip(ylim=c(0,60))

p2 <- ggplot(nrehousing, aes(x = rps)) +
        geom_histogram(colour="black", fill="white") +
        coord_cartesian(xlim=c(0,60)) 

grid.arrange(p1, p2, widths = c(1, 2))




##Conclusion
...

\pagebreak

##Task 3
...
```{r}


```

##Conclusion
...

