---
title: "QMB Exercise 2 - Estimation and Testing"
author: "Karin Gryzlak & Dino Nienhold"
date: "10 May, 2015"
output:
  pdf_document:
    fig_caption: yes
---

#Introduction
The following report is based on the QMB Exercise 2 - Estimation and Testing. The task description pdf file is bis_ex2-EstimationTesting-20150429.pdf

#Requirements
Please make sure that you the following packages loaded in your workspace.
```{r, message=F, warning=F}
library("dplyr")
library("ggplot2")
library("ggExtra")
library("gridExtra")
library("moments")
```

#Data Set
Please make sure you have the file housingrents.csv in the subdirectoy Data in your workspace.
```{r}
housingrents <- read.csv("./Data/housingrents.csv",sep=";")
```


##Data Processing
For analysis purposes it is necessary to convert the rooms and NRE variable to a factor. 
Furthermore a new variable rps (rent per square meter) is created.
Additionally the 2 sub datasets are created for NRE respectively non-NRE appartments.
```{r}
housingrents <- mutate(housingrents, rooms = factor(rooms), 
  nre = factor(nre,levels=c(0,1),labels=c("no","yes")))
housingrents <- mutate(housingrents,rps = rent/area)
nrehousing <- filter(housingrents,nre=="yes")
nonnrehousing <- filter(housingrents,nre=="no")
```

\pagebreak

##Task 1
In this task sample mean, standard error of the mean, conf interval, t-test and probablity is calculated.

###a)
For the dataset of 8 repair cases (2.6,12.2,8.3,28.6,0.5,19.0,16.3,5.7) the sample mean and standard error of the mean is calculated.

```{r}
x <- c(2.6, 12.2, 8.3, 28.6, 0.5, 19.0, 16.3, 5.7)
```

```{r, results=FALSE}
me <- mean(x)
n <- length(x)
se <- sd(x)/sqrt(n)
```
The mean is `r me`
and the standard error of the mean is `r se` 

###b)
In this sub task the normal plot is drawn for the dataset x. In figure 1 the points on this plot form a nearly linear pattern, which indicates that the normal distribution is a good model for this data set. 
```{r,fig.height=4 ,fig.cap="QQ Plot for dataset x"}
print(gg_qq(x))
```

###c)
In this sub task a 95% confidence interval is created:
```{r, eval=FALSE}
me + c(-1,1) * qnorm(.975) * se
```

We are 95% confident that the the interval from `r me + c(-1) * qnorm(.975) * se` to `r me + c(1) * qnorm(.975) * se` contains the mean service times of repair cases of this company.

###d)

A Student's t-test is conducted to see if the mean of x is less than four hours.

$\sf{H_{0}}$ = Mean is equal 4.

$\sf{H_{a}}$ = Mean is greater than 4.

```{r, echo=FALSE}
xTTestGreater <- t.test(housingrents$rps~housingrents$nre,alternative = "greater", mu=0, var.equal = FALSE)
```
The p-value is `r xTTestGreater$p.value` and much higher than 0.05. 
The mean difference is `r xTTestGreater$estimate[[2]]-xTTestGreater$estimate[[1]] ` and lies in the 95% confidence interval of the estimated population mean of `r xTTestGreater$conf.int[1]` and  Infinity. 
Therefore we fail to reject $\sf{H_{0}}$.

```{r}
t.test(x,mu=4, alternative = "greater")
```

###e)
In this sub task the probability the the repair time is larger than 24 hours is calculated.
```{r}
pnorm(24, mean(x), sd(x), lower.tail=FALSE)
```
The probability for a repair time of 24 hours or more is 9.40%. 

\pagebreak

##Task 2
Task 2 checks normality of variable rent per square (rps). Additionally t-tests are conducted

### a)
In this sub task the distribution of the variable rps for NRE respectively non-NRE appartments are checked for normality.
Figure 2 shows that the distribution is right skewed with several outliers on the higher rps range. 
Skewness value is `r skewness(nrehousing$rps)`.
Figure 3 shows a qq plot. The plot does also show the outliers and the non-normality of the distribution. 
```{r, fig.cap="Boxplot and Histogramm for variable rps for NRE appartments"}
p1 <- qplot(x = 1, y = rps, data = nrehousing, xlab = "", geom = 'boxplot') +
        coord_flip(ylim=c(0,60)) 
        

p2 <- ggplot(nrehousing, aes(x = rps)) +
        geom_histogram(colour="black", fill="white") +
        coord_cartesian(xlim=c(0,60)) +
        scale_y_continuous(breaks=seq(0,10,2))

grid.arrange(p1, p2, widths = c(1, 2))
```

\pagebreak

```{r, fig.cap="QQ Plot for variable rps for NRE appartments"}
print(gg_qq(nrehousing$rps))
```

\pagebreak
Check normal distribution of rps variable for non-NRE appartments:
Figure 4 shows that the distribution is right skewed with several outliers on the higher rps range. 
Skewness value is `r skewness(nonnrehousing$rps)`.
Figure 5 shows a qq plot. The plot does also show the outliers and the non-normality of the distribution.
```{r, fig.cap="Boxplot and Histogramm for variable rps for non-NRE appartments"}
p1 <- qplot(x = 1, y = rps, data = nonnrehousing, xlab = "", geom = 'boxplot') +
        coord_flip(ylim=c(0,60))

p2 <- ggplot(nonnrehousing, aes(x = rps)) +
        geom_histogram(colour="black", fill="white") +
        coord_cartesian(xlim=c(0,60)) +
        scale_y_continuous(breaks=seq(0,10,2))

grid.arrange(p1, p2, widths = c(1, 2))
```

\pagebreak

```{r, fig.cap="QQ Plot for variable rps for non-NRE appartments"}
print(gg_qq(nonnrehousing$rps))
```

\pagebreak

### b)
In this section a two-sided Student's t-test is conducted in order to check the following hypothesis:

$\sf{H_{0}}$ = Mean diference of the variable rps for NRE and non-NRE appartments is equal 0.

$\sf{H_{a}}$ = Mean diference of the variable rps for NRE and non-NRE appartmentsis not equal 0.

In the first t-test equal variance is not assumed. The p-value is `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = FALSE)$p.value` and much higher than 0.05. The mean difference is `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = FALSE)$estimate[[2]]-t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = FALSE)$estimate[[1]] ` and lies in the 95% confidence interval of the estimated population mean of `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = FALSE)$conf.int[1]` and `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = FALSE)$conf.int[2]` . Therefore we fail to reject $\sf{H_{0}}$.
```{r}
t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = FALSE)
```

In the second t-test equal variance is  assumed. The p-value is `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = TRUE)$p.value` and much higher than 0.05. The mean difference is `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = TRUE)$estimate[[2]]-t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = TRUE)$estimate[[1]] ` and lies in the 95% confidence interval of the estimated population mean of `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = TRUE)$conf.int[1]` and `r t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = TRUE)$conf.int[2]` . Therefore we fail to reject $\sf{H_{0}}$.
```{r}
t.test(housingrents$rps~housingrents$nre,alternative = "two.sided", mu=0, var.equal = TRUE)
```

\pagebreak

### c)
```{r,echo=FALSE,eval=TRUE}
rpsTTestGreater <- t.test(housingrents$rps~housingrents$nre,alternative = "greater", mu=0, var.equal = FALSE)
```
In this section a one-sided Student's t-test is conducted in order to check the following hypothesis:

$\sf{H_{0}}$ = Mean diference of the variable rps for NRE and non-NRE appartments is equal 0.

$\sf{H_{a}}$ = Mean diference of the variable rps for NRE and non-NRE appartments isgreater 0. 

In the t-test equal variance is  assumed. The p-value is `r rpsTTestGreater$p.value` and much higher than 0.05. 
The mean difference is `r rpsTTestGreater$estimate[[2]]-rpsTTestGreater$estimate[[1]] ` and lies in the 95% confidence interval of the estimated population mean of `r rpsTTestGreater$conf.int[1]` and  Infinity. 
Therefore we fail to reject $\sf{H_{0}}$.

```{r}
t.test(housingrents$rps~housingrents$nre,alternative = "greater", mu=0, var.equal = FALSE)
```
\pagebreak

##Task 3
In this task we conduct chi-square test of indepedence to test if the variable rooms is indepent from the variable nre.

###a)
In this subtask the chi-square test is conducted.
```{r}
housingrentTbl <- xtabs(~rooms+nre, data=housingrents)
```

```{r}
chisq.test(housingrentTbl) 

```

As the expected chi-square test value for the six room appartments is below 5. 
The chi-square test computation of p values is done by Monte Carlo simulation.
```{r}
chisq.test(housingrentTbl,simulate.p.value=TRUE) 
```

The p value 0.000499 is and below 0.01. 
Therefore we reject  $\sf{H_{0}}$ and there is a strong evidence that the variables are not independent.


###b)
The residuals for the chi-square test are evaluated in this sub task.
```{r}
resid(chisq.test(housingrentTbl,simulate.p.value=TRUE))
```
The 3, 4 and 5 room appartments belonging to NRE have the lowest respectively lowest residuals. This means that they have the highest impact on the test.  

\pagebreak

#Appendix

##Functions

Plot qqplot with ggplot
```{r,eval=FALSE}
gg_qq <- function(x, distribution = "norm", ..., line.estimate = NULL, conf = 0.95,
                  labels = names(x)){
        q.function <- eval(parse(text = paste0("q", distribution)))
        d.function <- eval(parse(text = paste0("d", distribution)))
        x <- na.omit(x)
        ord <- order(x)
        n <- length(x)
        P <- ppoints(length(x))
        df <- data.frame(ord.x = x[ord], z = q.function(P, ...))
        
        if(is.null(line.estimate)){
                Q.x <- quantile(df$ord.x, c(0.25, 0.75))
                Q.z <- q.function(c(0.25, 0.75), ...)
                b <- diff(Q.x)/diff(Q.z)
                coef <- c(Q.x[1] - b * Q.z[1], b)
        } else {
                coef <- coef(line.estimate(ord.x ~ z))
        }
        
        zz <- qnorm(1 - (1 - conf)/2)
        SE <- (coef[2]/d.function(df$z)) * sqrt(P * (1 - P)/n)
        fit.value <- coef[1] + coef[2] * df$z
        df$upper <- fit.value + zz * SE
        df$lower <- fit.value - zz * SE
        
        if(!is.null(labels)){ 
                df$label <- ifelse(df$ord.x > df$upper | 
                                           df$ord.x < df$lower, labels[ord],"")
        }
        
        p <- ggplot(df, aes(x=z, y=ord.x)) +
                geom_point() + 
                geom_abline(intercept = coef[1], slope = coef[2]) +
                geom_ribbon(aes(ymin = lower, ymax = upper), alpha=0.2) 
        if(!is.null(labels)) p <- p + geom_text( aes(label = label))
        return(p)
       }
```
