---
title: "QMB Exercise 2 - Estimation and Testing"
author: "Karin Gryzlak & Dino Nienhold"
date: "10 May, 2015"
output:
  pdf_document:
    fig_caption: yes
    
---

#Introduction
The following report is based on the QMB Exercise 2 - Estimation and Testing. The task description pdf file is bis_ex2-EstimationTesting-20150429.pdf

#Requirements
Please make sure that you the following packages loaded in your workspace.
```{r, message=F, warning=F}
library("dplyr")
library("ggplot2")
```

#Data Set
Please make sure you have the file housingrents.csv in the subdirectoy Data in your workspace.
```{r}
housingrents <- read.csv("./Data/housingrents.csv",sep=";")
```
\pagebreak

##Data Processing
For analysis purposes it is necessary to convert the rooms and NRE variable to a factor. Furthermore a new variable rps (rent per square meter) is created
```{r}
housingrents <- mutate(housingrents, rooms = factor(rooms), 
  nre = factor(nre,levels=c(0,1),labels=c("no","yes")))
housingrents <- mutate(housingrents,rps = rent/area)
```

##Task 1
...
```{r}

```

##Conclusion
...


\pagebreak

##Task 2
...
```{r}
The qqplot should help to get a better understanding wether the distribution of the variable is normally distributed.
```{r}
qqnorm(housingrents$rps)
qqline(housingrents$rps, col = 2)
```
```

##Conclusion
...

\pagebreak

##Task 3
...
```{r}


```

##Conclusion
...

